Data Exploration: IUCN and AquaMaps

### 2015-12-20 data_explore/data_explore.Rmd
cleaning up directory, separating manuscript into main and children

spp_list <- read.csv(data_file, stringsAsFactors = FALSE) %>%
  filter(!is.na(sm_perc)) 
  ### will ditch the ones where no IUCN polygon exists - or are these species
  ### whose ranges only occur in cells beyond the LOICZID grid?
spp_list <- spp_list %>%
  left_join(read.csv('sciname_lookups/spp_groupnames.csv', stringsAsFactors = FALSE),
            by = 'spp_group')

@@

sciname_lookup <- read.csv('sciname_lookups/spp_iucn_all.csv', stringsAsFactors = FALSE)

>>>>

spp_list <- read.csv(data_file, stringsAsFactors = FALSE) %>%
  filter(!is.na(sm_perc)) 
  ### will ditch the ones where no IUCN polygon exists - or are these species
  ### whose ranges only occur in cells beyond the LOICZID grid?
spp_list <- spp_list %>%
  left_join(read.csv('data/sciname_lookups/spp_groupnames.csv', stringsAsFactors = FALSE),
            by = 'spp_group')

@@

sciname_lookup <- read.csv('data/sciname_lookups/spp_iucn_all.csv', stringsAsFactors = FALSE)

### 2015-12-20 data_explore/data_explore_global_maps.Rmd
cleaning up directory, separating manuscript into main and children

@@

@@

levelplot(all_am_map, par.settings = myTheme, main = 'Entire Aquamaps Dataset')
```

@@

levelplot(all_iucn_map, par.settings = myTheme, main = 'Entire IUCN Dataset')
```

@@

### plot maps with distributions as percentages.  Should it be 0 - 1, or 
### 0 - max value?
levelplot(all_am_pct_map, par.settings = myTheme, 
          main = 'Aquamaps Dataset percent of max', 
          colorkey = list( at = seq(0, 1, length = 15)))
levelplot(all_iucn_pct_map, par.settings = myTheme, 
          main = 'IUCN Dataset percent of max', 
          colorkey = list(at = seq(0, 1, length = 15)))
### this is the difference one. Should it be min value to max value?
levelplot(diff_pct_map, par.settings = myTheme, 
          main = 'Difference: AquaMaps - IUCN by percent', 
          colorkey = list(at = seq(-1, 1, length = 15)))

>>>>

am_ln_map <- log(am_sp_map + 1)
iucn_ln_map <- log(iucn_sp_map + 1)

@@

levelplot(am_ln_map, par.settings = myTheme, 
          main = 'Aquamaps Dataset', 
          colorkey = list( at = seq(0, maxValue(am_ln_map), length = 15)))

levelplot(iucn_ln_map, par.settings = myTheme, 
          main = 'IUCN Dataset', 
          colorkey = list(at = seq(0, maxValue(iucn_ln_map), length = 15)))

@@

levelplot(all_am_map, par.settings = myTheme, main = 'Entire Aquamaps Dataset')

all_am_ln_map <- log(all_am_map + 1)
all_am_ln_map <- all_am_ln_map/maxValue(all_am_ln_map)
levelplot(all_am_ln_map, par.settings = myTheme, main = 'Entire Aquamaps Dataset: log-transformed relative species richness', 
          colorkey = list(at = seq(0, 1, length = 15)))

am_ln_pts <- as.data.frame(rasterToPoints(all_am_ln_map))
spp_richness_am_plot <- ggplot(am_ln_pts, aes(x = x, y = y)) +
  ggtheme_map + 
  geom_raster(aes(fill = layer), alpha = .7) +
  scale_fill_continuous(low = 'white', high = 'blue') +
  borders('world', color='gray40', fill='gray45', size = .1) +  # create a layer of borders
  scale_x_continuous(breaks = seq(-180, 180, by = 30), expand = c(0, 2)) +
  scale_y_continuous(breaks = seq( -90,  90, by = 30), expand = c(0, 2)) +
  labs(title = 'AquaMaps Dataset: log-transformed relative species richness', x = NULL, y = NULL, fill = NULL)

spp_richness_am_plot
ggsave(file.path(dir_fig, 'all_am_ln_map.png'), height = 4, width = 8)

@@

levelplot(all_iucn_map, par.settings = myTheme, main = 'Entire IUCN Dataset')

all_iucn_ln_map <- log(all_iucn_map + 1)
all_iucn_ln_map <- all_iucn_ln_map/maxValue(all_iucn_ln_map)
levelplot(all_iucn_ln_map, 
          par.settings = myTheme, 
          margin = list(axis = TRUE),
          main = 'Entire IUCN Dataset: log-transformed relative species richness', 
          colorkey = list(at = seq(0, 1, length = 15)))

iucn_ln_pts <- as.data.frame(rasterToPoints(all_iucn_ln_map))
spp_richness_iucn_plot <- ggplot(iucn_ln_pts, aes(x = x, y = y)) +
  ggtheme_map + 
  geom_raster(aes(fill = layer), alpha = .7) +
  scale_fill_continuous(low = 'white', high = 'blue') +
  borders('world', color='gray40', fill='gray45', size = .1) +  # create a layer of borders
  scale_x_continuous(breaks = seq(-180, 180, by = 30), expand = c(0, 2)) +
  scale_y_continuous(breaks = seq( -90,  90, by = 30), expand = c(0, 2)) +
  labs(title = 'IUCN Dataset: log-transformed relative species richness', x = NULL, y = NULL, fill = NULL)

spp_richness_iucn_plot
ggsave(file.path(dir_fig, 'all_iucn_ln_map.png'), height = 4, width = 8)

@@

### plot maps with distributions as percentages.  Should it be 0 - 1, or 
### 0 - max value?
levelplot(all_am_pct_map, par.settings = myTheme, 
          main = 'Aquamaps Dataset percent of max')
levelplot(all_iucn_pct_map, par.settings = myTheme, 
          main = 'IUCN Dataset percent of max')
### this is the difference one. Should it be min value to max value?
levelplot(diff_pct_map, par.settings = myTheme, 
          main = 'Difference: AquaMaps - IUCN by percent')

diff_ln_map <- all_am_ln_map - all_iucn_ln_map
levelplot(diff_ln_map, 
          par.settings = myTheme, 
          pretty = TRUE,
          margin = list(axis = TRUE),
          main = 'Difference in relative species richness (log-transformed): AquaMaps - IUCN', 
          colorkey = list(at = seq(-1, 1, length = 15)))
maxValue(all_am_ln_map)
minValue(all_am_ln_map)
maxValue(all_iucn_ln_map)
minValue(all_iucn_ln_map)
maxValue(diff_ln_map)
minValue(diff_ln_map)
diff_ln_pts <- as.data.frame(rasterToPoints(diff_ln_map))

spp_richness_diff_plot <- ggplot(diff_ln_pts, aes(x = x, y = y)) +
  ggtheme_map + 
  geom_raster(aes(fill = layer)) +
  scale_fill_gradient2(low = 'red', mid = 'white', high = 'blue', midpoint = 0) +
  borders('world', color='gray40', fill='gray45', size = .1) +  # create a layer of borders
  scale_x_continuous(breaks = seq(-180, 180, by = 30), expand = c(0, 2)) +
  scale_y_continuous(breaks = seq( -90,  90, by = 30), expand = c(0, 2)) +
  labs(title = 'Difference in relative species richness (log-transformed): AquaMaps - IUCN', x = NULL, y = NULL)

spp_richness_diff_plot
